Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-12-02 WSSSPE3/report/WSSSPE3_review.tex

\item \textbf{Louise Kellogg and Lorraine Hwang: \textit{Advancing Earth Science
through Best Practices in Open Source Software: Computational Infrastructure for
Geodynamics}} Lorraine Hwang presented experiences with CIG: the Computational
Infrastructure for Geodynamics, a community software with a worldwide user base.
Like others, their main goals include high usability, sustainability, and
reproducibility. In order to achieve these goals, communication channels have
been developed, such as mailing lists, wikis, workshops, hackathons, tutorials,
and webinars. In order to contribute to the infrastructure, codes have to adhere
to specified minimum standards, but ideals are also given to encourage reaching
the aforementioned goals. These include, e.g., the use of version control,
certain coding styles, the presence and nature of code tests and documentation,
and certain user workflows.

\item \textbf{Lorraine Hwang, Joe Dumit, Alison Fish, Louise Kellogg, Mackenzie
Smith, and Laura Soito: \textit{Software Attribution for Geoscience Applications
in the Computational Infrastructure for Geodynamics}} In a second talk, Lorraine
Hwang mentioned various ways to cite efforts within the framework, including
science papers, code papers, user manuals, and the CIG website. An analysis of
the resulting citations showed that $80\%$ of papers that use CIG codes mention
the code name, and about the same number includes a citation. Only about $20\%$
acknowledge CIG. Within the same sample of papers, about one fifth use an URL to
cite codes (including non-CIG codes), and only about one eighth specify the
version used. Compared to other codes, CIG seems to be much better cited. This,
in part, is attributed to suggested citation forms communicated to the
community, and runtime tools that help collecting a relevant citation list
automatically.

>>>>

\item \textbf{Louise Kellogg and Lorraine Hwang: \textit{Advancing Earth Science
through Best Practices in Open Source Software: Computational Infrastructure for
Geodynamics}} Lorraine Hwang presented experiences with CIG: the Computational
Infrastructure for Geodynamics, a community software with a worldwide user base.
Like others, their main goals include high usability, sustainability, and
reproducibility. In order to achieve these goals, communication channels have
been developed, such as mailing lists, wikis, workshops, hackathons, tutorials,
and webinars. In order to contribute to the infrastructure, codes must adhere
to specified minimum standards with the desire that all codes are working toward target standards.
%, but ideals are also given to encourage reaching the aforementioned goals. 
These include, e.g., the use of version control,
certain coding styles, the presence and nature of code tests and documentation,
and certain user workflows.

\item \textbf{Lorraine Hwang, Joe Dumit, Alison Fish, Louise Kellogg, Mackenzie
Smith, and Laura Soito: \textit{Software Attribution for Geoscience Applications
in the Computational Infrastructure for Geodynamics}} In a second talk, Lorraine
Hwang mentioned various ways to cite efforts within the framework, including
science papers, code papers, user manuals, and the CIG website. An analysis of
the resulting citations showed that $80\%$ of papers that use CIG codes mention
the code name, and about the same number includes a citation. Only about $20\%$
acknowledge CIG. Within the same sample of papers, about one fifth use an URL to
cite codes (including non-CIG codes), and only about one eighth specify the
version used. Compared to other codes, CIG seems to be much better cited. 
This in part is attribute to requiring that donated software provide a citable paper specified in the User Manual. 
The project is working on tools and methods to generate attribution information automatically.
%This,
%in part, is attributed to suggested citation forms communicated to the
%community, and runtime tools that help collecting a relevant citation list
%automatically.

### 2015-12-02 WSSSPE3/report/wg_appendix_software_credit.tex

Initial discussions focused on both various mechanisms for, and the
philosophical approach behind, crediting software in scientific papers. These
began with proposals for various ways to credit software (or other research
products including data) that contributed more significantly than a generic
citation, including
\begin{itemize}

@@

Although this working group's discussions at WSSSPE3 did not focus much on the
topic of tenure, the group recognized that there is no standard
policy---generally even within a single university---for software products to be
included in promotion and tenure dossiers. Thus, it may be difficult to
encourage valuing software contributions across the United States or United
Kingdom and globally; furthermore, stakeholders are typically not tenured and
thus may not be influential enough to change the status quo.

>>>>

Initial discussions focused on both various mechanisms for, and the
philosophical approach behind, crediting software in scientific papers. These
began with proposals for various ways to credit software (or other research
products including data) that contributed more significantly than a generic
citation, including:
\begin{itemize}

@@

Although this working group's discussions at WSSSPE3 did not focus much on the
topic of tenure and professional advancement, 
%we should recognized that this problem goes beyond faculty to recognizing the role of "Research Software Engineers"
the group recognized that there is no standard
policy---generally even within a single university---for software products to be
included in promotion and tenure dossiers. Thus, it may be difficult to
encourage valuing software contributions across the United States or United
Kingdom and globally; furthermore, stakeholders are typically not tenured and
thus may not be influential enough to change the status quo.
